---
title: Research and Extension Experiences for Undergraduates at K-State
author: Lucky Mehra
date: '2019-08-30'
slug: research-and-extension-experiences-for-undergraduates-at-k-state
categories: []
tags:
  - undergraduate
  - research
  - extension
  - teaching
  - mentoring
authors: []
header:
  caption: ''
  image: ''
  preview: yes
output:
  blogdown::html_page:
    toc: true
---

## Introduction

This summer, we hosted eight undergraduate students from all over the country for our first year of Research and Extension Experiences for Undergraduates (REEU). This program is funded under the USDA-NIFA's Education and Workforce Development initiative. The purpose of the program is to train undergraduate students in research and extension. We will be training 35 students over the course of 4 years (2019 to 2022).

Each student had an assigned mentor in either plant pathology, entomology, or agronomy departments. In addition to working with their mentors on their research or extension projects, the students also participated in professional development seminars, field trips, and FACT workshops. FACT stands for Food and Agriculture Cyberinformatics Tools. Our goal was to introduce students to a wide variety of topics related to modern plant health research; so that, they have enough data in hand to make a decision about what they would like to pursue for further studies.

Based on the results of student surveys, the program was able to meet its goals. The success of the program is reflected in the level of skills that students gained over the short 9-week period. See below for some of the areas where students improved their knowledge and skills. These are a few examples of questions that we asked students in an anonymous survey.

```{r import-survey-data, echo=FALSE, message=FALSE}
# install package manager `pacman`
if(!require(pacman)){
	install.packages("pacman")
}

# install and load packages needed for this document
pacman::p_load(tidyverse,
							 likert,
							 googlesheets,
							 rlang,
							 knitr,
							 kableExtra)

# read the sheet from google drive
REEU <- gs_title("student_assessment")

# post-survey data
post_reeu_data <- REEU %>% 
	gs_read(ws = 2) %>% 
	dplyr::select(-question_prefix) %>% 
  filter(question_number %in% c(11, 17, 6, 21, 20, 29, 31)) %>% 
  dplyr::select(-question_number)

# question labels
post_reeu_lookup <- REEU %>% 
	gs_read(ws = 2) %>% 
	dplyr::select(question_number, question_prefix, question_text) %>% 
	slice(1:40) %>% 
	mutate(question = paste(question_prefix, question_text))

# make the dataset wide
post_reeu_data_num <- post_reeu_data %>% 
	dplyr::select(-response_chr) %>% 
	filter(!is.na(response_num)) %>% 
	spread(key = question_text,
				 value = response_num) %>% 
	select(-student_num) %>% 
	mutate_all(round, 0) %>% 
	mutate_all(factor, levels = c("1", "2", "3", "4", "5"), ordered = TRUE) %>% 
	as.data.frame()
```


1. **`r post_reeu_lookup$question_prefix[1]`**

```{r post-survey-plot-1, echo=FALSE}
plot1 <- post_reeu_data_num %>% 
	select(1:2, 7) %>% 
	likert() 

plot(plot1)
```

Scale: `1` = No gain, `2` = A little gain, `3` = Moderate gain, `4` = Good gain, `5` = Great gain

All students had a `moderate` to `great` gain in understanding of everyday researh work, confidence to contribute to science, and conducting scientific observations.

2. **`r post_reeu_lookup$question_prefix[20]`**

```{r post-survey-plot-2, echo=FALSE}
plot2 <- post_reeu_data_num %>% 
	select(3, 5) %>% 
	likert() 

plot(plot2)
```

Scale: `1` = None, `2` = A little, `3` = Some, `4` = A fair amount, `5` = A great deal

Majority of students felt like a scientist and were engaged in real-world science research.

3. **`r post_reeu_lookup$question_prefix[30]`**

```{r post-survey-plot-3, echo=FALSE}
plot3 <- post_reeu_data_num %>% 
	select(4) %>% 
	likert() 

plot(plot3)
```

Scale: `1` = Not more likely, `2` = A little more likely, `3` = Somewhat more likely, `4` = Much more likely, `5` = Extremely more likely

At the end of the REEU, 83% students were `much more` to `extremely` likely to enroll in a graduate program in agricultural and related sciences.

## Activities

For a timeline of activities that we hosted over the summer, search `#KSUplanthealthREEU` hashtag on [twitter](twitter.com) or just [click here](https://twitter.com/hashtag/KSUplanthealthREEU?src=hashtag_click). Here are our interns celebrating after the successful resereach symposium:

```{r echo=FALSE}
blogdown::shortcode("tweet", '1154497716296310789')
```

## Student work

During last week of the program, the students participated in a poster symposium. The collection of student work can be [found here](https://krex.k-state.edu/dspace/handle/2097/40101).

Video explanation of one of the project by the student intern, Sam Whitten:

```{r echo=FALSE}
blogdown::shortcode("youtube", "EcvNRaTBgW0")
```

We were also in K-State Research and Extension News [Students take on important plant-based research](https://www.ksre.k-state.edu/news/stories/2019/06/plant-pathology-NIFA-summer-internships.html)

## Application information for 2020

We will start accepting applications for summer 2020 in December 2019. Interested undergraduate students should watch [our REEU website](https://www.plantpath.k-state.edu/undergraduate/reeu/index.html) for application related updates.

## Program personnel

Please feel free to contact any three of us if you have any question about the program and the application process.

Director: [Megan Kennelly](https://www.plantpath.k-state.edu/people/faculty/kennelly/)

Coordinator and co-director: [Lucky Mehra](https://www.plantpath.k-state.edu/people/faculty/mehra/)

Co-director: [David Cook](https://www.plantpath.k-state.edu/people/faculty/cook-david/)